### 2019-09-12 ic-journal/multivariate.tex
Stage work on multivariate section

\section{The multivariate formalisation}
\label{sec:multivariate}

In this section, the formalisation of the multivariate (general)
cases of Milner's ``unique solution of equations (for $\sim$)''
(Theorem~\ref{t:Mil89s1}) and our ``unique solution of rooted
contractions'' (Theorem~\ref{t:rcontraBisimulationU}) is
described.
We choose these two theorems because they both depend on the same
concept of ``weakly guarded contexts'', and Lemma~\ref{lem:milner313}.
\hl{This work is entirely new} after~\cite{EPTCS276.10}.
It is accumulative, completely based on the
existing formal framework in the sense that, neither the CCS datatype nor
the SOS rules need to be changed.

The central problem of the multivariate formalisation is the
representation of CCS equations (expressions, contexts).
Recall the univariate formalisation, $\lambda$-functions (of type
$CCS\rightarrow CCS$) are used to represent univariate CCS
equations, and variable substitutions are simply done by
calling $\lambda$-functions on the target CCS terms.
This idea cannot be extended to the multivariate case, as
we don't have a fixed number of variables to deal with.

@@

\subsection{Multivariate substitutions}

There are two ways to represent (and actually do) the multivariate
substitution of $E\{\til P/\til X\}$ which substitutes each possibly
occurrences of free variables $X_i$ in $E$ with the corresponding
$P_i$: (1) iterately applying the existing univariate
\HOLinline{\HOLConst{CCS_Subst}}; (2) defining a new multivariate substitution
function. In any case, we (reasonably) assume that all $X_i$ are
distinct with each other. On the other hand, there is the syntactic
possibility that $\til P$ again contains variables from $\til X$, thus
different orders of iterated substitutions may lead to different results.
In fact, there are all kinds of issues with
substitutions and applying them to term-like structures. To obtain the
maximal flexibility with mimimal efforts, we defined
a multivariate version of \HOLinline{\HOLConst{CCS_Subst}}
called \HOLinline{\HOLConst{CCS_SUBST}} based HOL's \texttt{finite\_map} theory~\cite{holdesc}.

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Notice that, the variable substitution only happens on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must further does the
substitution on $E$ using a reduce map with $X$, so that all
occurences of agent variable $X$ in $E$ are correctly bypassed. On the
other hand, since \HOLinline{\HOLConst{CCS_SUBST}} only walks each subterm
once, even \HOLinline{\HOLFreeVar{Ps}} contains free variables, these variables will
not be further substituted. Thus the process is indeed
order-independent. Below we give some key lemmas about \HOLinline{\HOLConst{CCS_SUBST}}.

\begin{lemma}
  If there is only one single variable $X$ in the map (with the target
  expression $E'$), then \HOLinline{\HOLConst{CCS_SUBST}} behaves exactly the
  same as (the univariate) \HOLinline{\HOLConst{CCS_Subst}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CCS_SUBST} \ensuremath{(}\HOLConst{FEMPTY} \HOLSymConst{|+} \ensuremath{(}\HOLFreeVar{X}\HOLSymConst{,}\HOLFreeVar{E\sp{\prime}}\ensuremath{)}\ensuremath{)} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{CCS_Subst} \HOLFreeVar{E} \HOLFreeVar{E\sp{\prime}} \HOLFreeVar{X}\hfill{[CCS_SUBST_SING]}
\end{alltt}
\end{lemma}

\begin{lemma}
If the free variables of $E$ is disjoint with \HOLinline{\HOLFreeVar{Xs}}, a
substitution of \HOLinline{\HOLFreeVar{Xs}} in $E$ does not change $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{E}\hfill{[CCS_SUBST_elim]}
\end{alltt}
\end{lemma}

\begin{lemma}
Substituting each free variable $X$ in $E$ to \HOLinline{\HOLConst{var}\\\;\HOLFreeVar{X}} (itself)
does not change $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLConst{MAP} \HOLConst{var} \HOLFreeVar{Xs}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{E}\hfill{[CCS_SUBST_self]}
\end{alltt}
\end{lemma}

The next lemma plays an important role. It essentially swaps the order of two substitutions:
\begin{lemma}
Under certain conditions, two nested substitutions can be
converted into a single substitution where the targets are substituted first:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLConst{MAP} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\hfill{[CCS_SUBST_nested]}
\end{alltt}
\end{lemma}

Finally, the following two lemmas precisely estimate the free and
bound variables of a substituted term:
\begin{lemma}
The bound variables of \HOLinline{\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]}\\\;\HOLFreeVar{E}} have
at most the union of all bound variables from $E$ and \HOLinline{\HOLFreeVar{Ps}}.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLFreeVar{E} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{BV} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Ps}\ensuremath{)}\ensuremath{)}
\end{alltt}
\end{lemma}

\begin{lemma}
The free variables of \HOLinline{\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]}\\\;\HOLFreeVar{E}} have
at most the union of all free variables from $E$ and \HOLinline{\HOLFreeVar{Ps}}
removing \HOLinline{\HOLFreeVar{Xs}}.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \HOLConst{FV} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLFreeVar{E} \HOLConst{DIFF} \HOLConst{set} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{FV} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Ps}\ensuremath{)}\ensuremath{)}
\end{alltt}
\end{lemma}

It is worth noting that, in some of the above lemmas, the condition
\HOLinline{\HOLConst{DISJOINT}\\\;\ensuremath{(}\HOLConst{BV}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{set}\\\;\HOLFreeVar{Xs}\ensuremath{)}} (the disjointness between bound variables of $E$ and
 substitution variables) is not necessary but makes some proofs
 easier.

\subsection{Multivariate (weakly-guarded) contexts}

\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenDefEquality{}}
     \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{E}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}\ensuremath{)}

\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLConst{context} \HOLFreeVar{Xs}\ensuremath{)} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}}
   \HOLConst{context} \HOLFreeVar{Xs} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)}

@@

\subsection{Multivariate equations and solutions}

>>>>

\section{The \multivariate formalisation}
\label{sec:multivariate}

In this section, the formalisation of the \multivariate (general)
cases of Milner's ``unique solution of equations (for $\sim$)''
(Theorem~\ref{t:Mil89s1}) and our ``unique solution of rooted
contractions'' (Theorem~\ref{t:rcontraBisimulationU}) is
described.
We choose these two theorems because they both depend on the same
concept of ``weakly guarded contexts'', and Lemma~\ref{lem:milner313}.
\hl{This work is entirely new} after~\cite{EPTCS276.10}.
It is accumulative, completely based on the
existing formal framework in the sense that, neither the CCS datatype nor
the SOS rules need to be changed.

The central problem of the \multivariate formalisation is the
representation of CCS equations (expressions, contexts).
Recall the univariate formalisation, $\lambda$-functions (of type
$CCS\rightarrow CCS$) are used to represent univariate CCS
equations, and variable substitutions are simply done by
calling $\lambda$-functions on the target CCS terms.
This idea cannot be extended to the \multivariate case, as
we don't have a fixed number of variables to deal with.

@@

\subsection{\Multivariate substitutions}

There are two ways to represent (and actually do) the \multivariate
substitution of $E\{\til P/\til X\}$ which substitutes each possibly
occurrences of free variables $X_i$ in $E$ with the corresponding
$P_i$: (1) iterately applying the existing univariate
\HOLinline{\HOLConst{CCS_Subst}}; (2) defining a new \multivariate substitution
function. In any case, we (reasonably) assume that all $X_i$ are
distinct with each other. On the other hand, there is the syntactic
possibility that $\til P$ again contains variables from $\til X$, thus
different orders of iterated substitutions may lead to different results.
In fact, there are all kinds of issues with
substitutions and applying them to term-like structures. To obtain the
maximal flexibility with mimimal efforts, we defined
a \multivariate version of \HOLinline{\HOLConst{CCS_Subst}}
called \HOLinline{\HOLConst{CCS_SUBST}} based HOL's \texttt{finite\_map} theory~\cite{holdesc}.

@@

   \HOLConst{CCS_SUBST} \HOLFreeVar{fm} \ensuremath{(}\HOLConst{rec} \HOLFreeVar{X} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLKeyword{if} \HOLFreeVar{X} \HOLSymConst{\HOLTokenIn{}} \HOLConst{FDOM} \HOLFreeVar{fm} \HOLKeyword{then} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \ensuremath{(}\HOLFreeVar{fm} \HOLSymConst{\ensuremath{\setminus}} \HOLFreeVar{X}\ensuremath{)} \HOLFreeVar{E}\ensuremath{)}
     \HOLKeyword{else} \HOLConst{rec} \HOLFreeVar{X} \ensuremath{(}\HOLConst{CCS_SUBST} \HOLFreeVar{fm} \HOLFreeVar{E}\ensuremath{)}\hfill{[CCS_SUBST_rec]}
\end{alltt}
Notice that, the variable substitution only happens on free
variables. In the case of $\recu X E$: if $X$ (a bound variable) is in
the domain of \HOLinline{\HOLFreeVar{fm}}, \HOLinline{\HOLConst{CCS_SUBST}} must further does the
substitution on $E$ using a reduce map with $X$, so that all
occurences of agent variable $X$ in $E$ are correctly bypassed. On the
other hand, since \HOLinline{\HOLConst{CCS_SUBST}} only walks each subterm
once, even \HOLinline{\HOLFreeVar{Ps}} contains free variables, these variables will
not be further substituted. Thus the process is indeed
order-independent. Below we present some key lemmas about
\HOLinline{\HOLConst{CCS_SUBST}} \emph{without proofs}:

\begin{lemma}[\texttt{CCS_SUBST_SING}]
  If there is only one single variable $X$ in the map (with the target
  expression $E'$), then \HOLinline{\HOLConst{CCS_SUBST}} behaves exactly the
  same as (the univariate) \HOLinline{\HOLConst{CCS_Subst}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CCS_SUBST} \ensuremath{(}\HOLConst{FEMPTY} \HOLSymConst{|+} \ensuremath{(}\HOLFreeVar{X}\HOLSymConst{,}\HOLFreeVar{E\sp{\prime}}\ensuremath{)}\ensuremath{)} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{CCS_Subst} \HOLFreeVar{E} \HOLFreeVar{E\sp{\prime}} \HOLFreeVar{X}
\end{alltt}
\end{lemma}

\begin{lemma}[\texttt{CCS_SUBST_elim}]
If the free variables of $E$ is disjoint with \HOLinline{\HOLFreeVar{Xs}}, a
substitution of \HOLinline{\HOLFreeVar{Xs}} in $E$ does not change $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{FV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{E}
\end{alltt}
\end{lemma}

\begin{lemma}[\texttt{CCS_SUBST_self}]
Substituting each free variable $X$ in $E$ to \HOLinline{\HOLConst{var}\\\;\HOLFreeVar{X}} (itself)
does not change $E$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLConst{MAP} \HOLConst{var} \HOLFreeVar{Xs}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLFreeVar{E}
\end{alltt}
\end{lemma}

The next lemma plays an important role. It essentially swaps the order of two substitutions:
\begin{lemma}[\texttt{CCS_SUBST_nested}]
Under certain conditions, two nested substitutions can be
converted into a single substitution where the targets are substituted first:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{C}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLConst{MAP} \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}
\end{alltt}
\end{lemma}

Finally, the following two lemmas precisely estimate the free and
bound variables of a substituted term:
\begin{lemma}[\texttt{BV_SUBSET_BIGUNION}]
The bound variables of \HOLinline{\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]}\\\;\HOLFreeVar{E}} have
at most the union of all bound variables from $E$ and \HOLinline{\HOLFreeVar{Ps}}.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \HOLConst{BV} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{BV} \HOLFreeVar{E} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{BV} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Ps}\ensuremath{)}\ensuremath{)}
\end{alltt}
\end{lemma}

\begin{lemma}[\texttt{FV_SUBSET_BIGUNION_PRO}]
The free variables of \HOLinline{\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]}\\\;\HOLFreeVar{E}} have
at most the union of all free variables from $E$ and \HOLinline{\HOLFreeVar{Ps}}
removing \HOLinline{\HOLFreeVar{Xs}}.
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \HOLConst{FV} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{FV} \HOLFreeVar{E} \HOLConst{DIFF} \HOLConst{set} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenUnion{}} \HOLConst{BIGUNION} \ensuremath{(}\HOLConst{IMAGE} \HOLConst{FV} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Ps}\ensuremath{)}\ensuremath{)}
\end{alltt}
\end{lemma}

It is worth noting that, in the above lemmas, the condition
\HOLinline{\HOLConst{DISJOINT}\\\;\ensuremath{(}\HOLConst{BV}\\\;\HOLFreeVar{E}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{set}\\\;\HOLFreeVar{Xs}\ensuremath{)}} (the bound
variables of $E$ and the free substitution variables are disjoint) is
not necessary but makes some proofs (much) easier. As these lemmas are
used in the rest work, these disjointness conditions eventually
are propagated everywhere. (We believe that they can be
all eliminated in the future.)

\subsection{\Multivariate (weakly-guarded) contexts}

Recall in Section~\ref{ss:context}, univariate contexts and their
guarded companions are defined as predicates over $\lambda$-expressions
of type $CCS\rightarrow CCS$. These predicates, such as
\HOLinline{\HOLConst{CONTEXT}}
(contexts/expressions), \HOLinline{\HOLConst{WG}} (weak-guarded contexts),
\HOLinline{\HOLConst{SG}} (strongly-guarded contexts) and \HOLinline{\HOLConst{SEQ}} (sequential
contexts), are all defined inductively, i.e.~they are built from some ``holes''
in a bottom-up way. For instance, \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{\mid}}\\\;\HOLFreeVar{P}\ensuremath{)}}
holds because \HOLinline{\HOLConst{WG}\\\;\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}} holds on the ground,
and for any agent variable $X$, we have (by $\beta$-reduction)
\HOLinline{\ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}.\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}\ensuremath{)}\\\;\ensuremath{(}\HOLConst{var}\\\;\HOLFreeVar{X}\ensuremath{)}\\\;\HOLSymConst{\ensuremath{=}}\\\;\HOLFreeVar{l}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{P}} (or ``$l.X + P$'' in textbook
notation) is weakly guarded (c.f.~Def.~\ref{def:guardness}).

It is possible to inductively assert multivariate contexts and their
guarded variants, but this is not easy as we need to carefully exclude the cases where
free variables occur inside any recusion operator (c.f. the footnote
at the beginning of Section~\ref{ss:SysEq}. A more elegant (and less
error-prone) solution is to define \multivariate contexts (and variants) upon the
existing \univariate definitions. This is indeed our choice.

For instance, to see the weak guardedness of $a.X + b.X + c.Y$, it is
natural to focus on each (free) variable: if we substitute each
occurrence of just one variable, say $X$, with a ``hole'' and see the
resulting term as a $\lambda$-function, then all $\lambda$-functions
derived from the original expression in this way must satisfy
\HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}
\end{alltt}
Notice that, \HOLinline{\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\\\;\HOLSymConst{\ensuremath{=}}\\\;\ensuremath{[}\HOLFreeVar{t}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\ensuremath{(}\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\ensuremath{)}}. This idea leads to the following definitions:
\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}

   \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}
\end{alltt}

\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{C} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLConst{context} \HOLFreeVar{Xs}\ensuremath{)} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}}
   \HOLConst{context} \HOLFreeVar{Xs} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Es}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{C}\ensuremath{)}

@@

\subsection{\Multivariate equations and solutions}
